---
title: "p8105_hw6_zl3119"
author: "Zheyan"
date: "11/24/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)
library(fastDummies)
library(glmnet)

set.seed(777)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = 'bottom'))

options(
  ggplot2.continuous.colour = 'viridis',
  ggplot2.continuous.fill = 'viridis'
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

In this problem, you will analyze data gathered to understand the effects of several variables on a child’s birthweight. This dataset, available here, consists of roughly 4000 children and includes the following variables:

## Load and clean the data for regression analysis

```{r read and clean data}
birthweight_df = 
  read_csv('data/birthweight.csv', show_col_types = FALSE) %>% 
  janitor::clean_names()

# No missing value
colSums(is.na(birthweight_df))

birthweight_df = 
  birthweight_df %>% 
  # change some numeric variables to factors
  mutate(babysex = as.factor(babysex),
         # rename as father_race
         father_race = as.factor(frace),
         malform = as.factor(malform),
         # rename as mother_race
         mother_race = as.factor(mrace),
         if_smoke = as.factor(ifelse(smoken > 0, 1, 0))) %>% 
  select(-frace, -mrace, -smoken)

# show the cleaned data
head(birthweight_df) %>% 
  knitr::kable()
```

There isn't missing value in the dataframe. Change 'babysex', 'father_race', 'malform', 'mother_race' into factor. In addition, change smoken into an indicator variable 'if_smoke' because the average number of cigarettes smoked per day during pregnancy does not have linear association with the weight gain of the baby.


## Build model

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values.

Use Lasso to select variables
```{r pressure, echo=FALSE}

# dummy variables
df_dummy = dummy_cols(birthweight_df, 
                      select_columns = c('babysex', 'father_race', 'malform', 'mother_race', 'if_smoke'),
                      remove_selected_columns = TRUE,
                      remove_first_dummy = TRUE)


#perform k-fold cross-validation to find optimal lambda value
x = data.matrix(
      df_dummy %>% 
      select(-bwt)
)

y = data.matrix(
      df_dummy %>% 
      pull(bwt)
)

cv_model = cv.glmnet(x, y, alpha = 1)

#find optimal lambda value that minimizes test MSE
best_lambda = cv_model$lambda.1se


#produce plot of test MSE by lambda value
plot(cv_model) 

best_model = glmnet(x, y, alpha = 1, lambda = best_lambda)
coef.apprx = coef(best_model, s = 0.5, exact = FALSE)
coef(best_model) 


# coef.apprx[which(coef.apprx != 0)]
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
