---
title: "p8105_hw6_zl3119"
author: "Zheyan"
date: "11/24/2021"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(modelr)
library(mgcv)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = 'bottom'))

options(
  ggplot2.continuous.colour = 'viridis',
  ggplot2.continuous.fill = 'viridis'
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

knitr::opts_chunk$set(echo = TRUE)
```

# Problem 1

In this problem, you will analyze data gathered to understand the effects of several variables on a child’s birthweight. This dataset, available here, consists of roughly 4000 children and includes the following variables:

## Load and clean the data for regression analysis

```{r cars}
birthweight_df = 
  read_csv('data/birthweight.csv') %>% 
  janitor::clean_names()

# No missing value
colSums(is.na(birthweight_df))

birthweight_df = 
  birthweight_df %>% 
  # change some numeric variables to factors
  mutate(babysex = as.factor(babysex),
         # rename as father_race
         father_race = as.factor(frace),
         malform = as.factor(malform),
         # rename as mother_race
         mother_race = as.factor(mrace),
         smoken = as.factor(smoken)) %>% 
  select(-frace, -mrace)

birthweight_df %>% 
  knitr::kable()
```

## Build model

Propose a regression model for birthweight. This model may be based on a hypothesized structure for the factors that underly birthweight, on a data-driven model-building process, or a combination of the two. Describe your modeling process and show a plot of model residuals against fitted values.


```{r pressure, echo=FALSE}
library(glmnet)

#perform k-fold cross-validation to find optimal lambda value
x = data.matrix(
      birthweight_df %>% 
      select(-bwt)
)

y = data.matrix(
      birthweight_df %>% 
      pull(bwt)
)

cv_model <- cv.glmnet(x, y, alpha = 1)

#find optimal lambda value that minimizes test MSE
best_lambda <- cv_model$lambda.min
best_lambda


#produce plot of test MSE by lambda value
plot(cv_model) 
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
